---
title: "run_analysis.R"
author: "James Chen"
date: "Friday, July 25, 2014"
output: html_document
---

Due to campatibility issue this README.Rmd can't be run with Knit HTML. 
However, it can be serves as a documentation for the source run_analysis.R code.

Requirement:
   
Create one R script called run_analysis.R that does the following. 
 
Merges the training and the test sets to create one data set.
Extracts only the measurements on the mean and standard deviation for each measurement. 
Uses descriptive activity names to name the activities in the data set
Appropriately labels the data set with descriptive variable names. 
Creates a second, independent tidy data set with the average of each variable for each activity and each subject. 


Setting Directory for Downlaoding and Unzip file
```{r}
ProjectDirectory = getwd()
DataDirectory = "UCI HAR Dataset/"
dataFile = "dataset.RData"
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", 
               "data.zip", quiet = TRUE, mode = "wb")  
unzip("data.zip")
```

Reading labels
```{r}
  setwd(DataDirectory)
  
  temp = read.table("activity_labels.txt", sep = "") 
  activityLabels = as.character(temp$V2)
```

Read feature names
```{r}
  temp = read.table("features.txt", sep = "")
  attributeNames = temp$V2
```
  
Read X_train data
```{r}
  Xtrain = read.table("train/X_train.txt", sep = "")
```

Setting attribute names
```{r}
names(Xtrain) = attributeNames

```
  
Read y_train data
```{r}
  Ytrain = read.table("train/y_train.txt", sep = "")
```

Settting names
```{r}
  names(Ytrain) = "Activity"
```

Factorization
```{r}
Ytrain$Activity = as.factor(Ytrain$Activity)
```

Requirement: Uses descriptive activity names to name the activities in the data set
Provides access to the levels attribute of a variable

```{r}
  levels(Ytrain$Activity) = activityLabels
```
  
Read Subject data
```{r}
  trainSubjects = read.table("train/subject_train.txt", sep = "")
  names(trainSubjects) = "subject"
```

Setting Subject as factor
```{r}
  trainSubjects$subject = as.factor(trainSubjects$subject)
```

Combing data
```{r}
  train = cbind(Xtrain, trainSubjects, Ytrain)
```

Read test data same as train data

```{r}

  Xtest = read.table("test/X_test.txt", sep = "")
  names(Xtest) = attributeNames

  Ytest = read.table("test/y_test.txt", sep = "")
  names(Ytest) = "Activity"

  Ytest$Activity = as.factor(Ytest$Activity)
  levels(Ytest$Activity) = activityLabels
  
  testSubjects = read.table("test/subject_test.txt", sep = "")
  names(testSubjects) = "subject"

  testSubjects$subject = as.factor(testSubjects$subject)
  test = cbind(Xtest, testSubjects, Ytest)
```

Save to data file
```{r}
  save(train, test, file = dataFile)
```
Remove objects for saving memoary
```{r}
  rm(train, test, temp, Ytrain, Ytest, Xtrain, Xtest, trainSubjects, testSubjects, 
     activityLabels, attributeNames)
```


Reload objects
```{r}
load(dataFile)
```

Setting working directory
```{r}
setwd(ProjectDirectory)
```

Adding new columns
```{r}
train$Partition = "Train"
test$Partition = "Test"
```
Merges the training and the test sets to create one data set.
```{r}
all = rbind(train, test)  # combine sets for visualization
all$Partition = as.factor(all$Partition)
```


Fielter then Combine
```{r}
library(plyr)
library(reshape)

cindex <-grepl(pattern = "mean\\(|std\\(", names(all), perl=TRUE)

cindex[ (length(cindex)-2):  (length(cindex)) ] <- TRUE
extract<-all[, cindex]
```

Save object fo a data file
```{r}
save(train, test, extract, file = dataFile)

```

Reload the data file
```{r}
load(dataFile)
setwd(ProjectDirectory)
```

Melt an object into a form suitable for easy casting.
```{r}
tidydata <- melt(extract,id=67:69,measure.vars=1:66)
```


Create final tidy data
```{r}
newtidydata<-ddply(tidydata,.(subject,Activity,Partition,variable),summarize,value=mean(value))
```

Write to a file
```{r}
write.table(newtidydata, "mytidydata.txt", sep="\t") 
```

